An Introduction to Rocker:\\ Docker Containers for R

### 2017-11-23 boettiger-eddelbuettel.tex
'pinning' rewrite
(the diff shows changes in .tex too which may be editor-whitespace
stuff; I only change the 'debian testing' section as the diff to Rmd shows)

Docker is a popular open-source tool to create, distribute, deploy, and
run software applications using \emph{containers}. Containers provide a
virtual environment (see \citet{Clark2014} for an overview of common
virtual environments) requiring all operating-system components an
application needs to run: code, runtime, system tools, system runtime.
Docker containers are lightweight as they share the operating system
kernel, starting instantly using a layered filesystem which minimizes
disk footprint and download time, are built on open standards that run
on all major platforms (Linux, Mac, Windows), and provide an added layer
of security by running an application in an isolated environment
\citep{what-docker}. Familiarity with a few key terms is helpful in
understanding this paper. The term ``container'' refers to an isolated
software environment on a computer. R users can think of running a
container as analogous to loading an R package; it makes a set of
software functions available. A Docker ``image'' is a binary archive of
that software, analogous to an R binary package: a given version is
downloaded only once, and can then be ``run'' to create a container
whenever it is needed. A ``Dockerfile'' is a recipe, the source-code, to
create a Docker image. Development and contributions to the Rocker
project focuses on the construction, organization and maintenance of
these Dockerfiles.

\section{Design principles \& use
cases}\label{design-principles-use-cases}

Docker gives users very convenient access to pre-configured and
pre-built \emph{binary} images that ``just work''. This allows R users
to access a wider-variety of ready-to-use environments than provided by
either the R Project itself or, say, their distribution which will
generally focus on one (current) release. For example, R users on
Windows may run RStudio Server or Shiny Server locally just by launching
a single command (once Docker itself is installed). Another common
use-case is access to R-devel without affecting the local system. Here,
we detail some of the principle use cases motivating these containerized
versions of R environments, and the design principles that help make
them work.

@@

One common use case for Rocker containers is to provide a fast and
reliable mechanism to deploy a custom R environment to a remote server,
such as Amazon Web Services Elastic Compute (AWS EC2), DigitalOcean,
NSF's Jetstream servers \citep{jetstream}, or private or institutional
server hardware. Rocker containers are also easy to run locally on most
modern laptops by first installing the appropriate Docker command (or
distribution) for Windows, MacOS, and Linux-based operating systems. By
sharing volumes with the local host, users can still manipulate files
with familiar, native tools while performing computation through a
reproducible, containerized environment \citep{Boettiger2015}. Being
able to test code in a predictable, pre-configured R environment on a
local machine and to then run the same code in an identical environment
on a remote server (\emph{e.g.}, for access to greater RAM, more
processors, or merely to free up the local machine from a long-running
computation) is essential for low-friction scaling of analysis. Without
such containerization, getting code to run appropriately in a remote
environment can be a major undertaking, requiring both time and
knowledge many would-be users may not have.

For instance, the following bash commands install Docker on almost any
Linux-based server and then launch a Rocker container providing the
RStudio-server environment over a web interface.

@@

The \texttt{docker\ run} option \texttt{-p} sets the port on which
RStudio will appear, 8787 is the RStudio default (adding your user to
the docker group to avoids the need for a \texttt{sudo} command to call
\texttt{docker}: \texttt{sudo\ usermod\ -g\ docker\ \$USER}). Many
academic and commercial cloud providers make it possible to execute such
code snippets when a container is launched, without ever needing to
\texttt{ssh} into the machine. The user may log into the server merely
by pasting its IP address or DNS name (followed by the chosen port,
\emph{e.g.}, \texttt{:8787}) into a browser and entering the appropriate
password. This provides the user with a familiar, interactive
environment running on a remote machine while requiring a minimum of
expertise.

Deploying Docker containers on centrally administrated multi-user
machines, such as department or university clusters, has previously been
more difficult, since system adminstrators do not want to allow the
elevated user permissions the Docker runtime environment requires. To
work around this problem, Lawrence Berkeley National Labs (LBNL) has
made `Singularity' \citep{singularity}: a container runtime environment
that users can both install and use to run most Docker containers
without requiring root privileges, making it easier to deploy Rocker
containers in this context as well.

@@

@@

Users can access a root \texttt{bash} shell in a running Rocker
container using

@@

As with any interactive Docker container, users should specify the
interactive (\texttt{-i}) and terminal (\texttt{-t}, here combined with
interactive as \texttt{-ti}) flags, and specify the desired executable
environment (\emph{e.g.}, \texttt{R}, though other common options may be
\texttt{Rscript} or \texttt{bash}):

@@

Another feature of Rocker containers is the ability to provide a
sandboxed environment, isolated from software and potentially from other
data on the machine. Many users are reluctant to upgrade their suite of
installed packages, which may break their existing code or even their R
environment if the installation goes poorly. However, upgrading packages
and/or the R environment is often necessary to run analysis of a
colleague, or access more recent methods. Rocker offers an easy
solution. For instance, a user can run R code requiring the most recent
versions of R and related packages inside a Rocker container without
having to upgrade their local installations first. Conversely, one could
use Rocker to run code on an older R release with prior versions of R
packages, again without having to make any alteration to one's local R
install. Another common use case is to access a container with support
for particular options such as using gcc or clang compiler sanitizers
\citep{edd_sanitizers}. These requires R itself be built with
specialized settings that may not be not available or familiar to many R
users on their native system, but can be easily deployed by pulling the
Rocker images \texttt{rocker/r-devel-san} or
\texttt{rocker/r-devel-ubsan-clang}.

This sandboxing feature is also valuable in the remote computing
context, allowing system administrators to grant users freedom to
install software which requires root privileges inside a container,
while not granting them root access on the host machine. Root access is
required to launch Docker containers, though not to access containers
already running and providing some service such as RStudio. Users
logging into a container through the
RStudio\textsuperscript{\textregistered} interface do not by default
have root privileges, though are able to install R packages. Granting
these users root privileges in the container still leaves them sandboxed
from the host container. Further, Docker images can be deployed from a
machine without root privileges through the alternate container runtime
environment, Singularity \citep{singularity}. Unlike traditional virtual
machines, these containers do not impose a large footprint of reserved
resources as a typical host can easily support 100s of containers
\citep{what-docker}.

@@

Users can easily determine the software stack installed on any Rocker
image by examining the associated Dockerfile recipe, which provides a
concise, human-readable record of the installation. All Rocker images
use automated builds through Docker Hub, which also acts as the central,
default repository distributing the images. Using automated builds
rather than uploading pre-built image binaries to the Docker Hub avoids
the potential for the build not to match the recipe. The corresponding
Dockerfile is visible both on the Docker Hub and in the linked GitHub
repository of the Rocker project (\url{https://github.com/rocker-org}),
which provides a transparent versioned history of all changes made to
these recipes, as well as documentation, a community wiki, and issue
trackers for discussing proposed changes, bugs, improvements to the
Dockerfiles and troubleshoot any issues users may encounter. Having
these public source files built automatically by a trusted provider
(Docker Hub), rather than built locally and uploaded only in binary form
provides is also useful from a security perspective in avoiding malware.

@@

Having a shared, transparent computational environment created by a
publicly hosted, reproducible recipe facilitates community input into
configuration details. R and many of its packages and related software
can be configured with a wide range of options, compilers, different
linear-algebra libraries and so forth. While this flexibility reflects
varying needs, many users rely on default settings which may be
optimized more for simplicity of installation than performance. The
Rocker recipes reflect significant community input on these choices, as
well as the considerable experience and expertise of the R Debian
maintainer of 20 years in what configuration options to use. This helps
create a more finely tuned, optimized reference implementation of the R
environment as well as a platform for comparing and discussing these
concerns which are often overlooked elsewhere. Issues and Pull Requests
on the Rocker repositories on GitHub attest to these discussions and
improvements. Widespread use of the Rocker image helps promote both
testing of these choices and contributions further tweaking the
configuration from many members of the R community.

@@

To make extensions transparent and persistent, Rocker images can be
extended by any user by writing their own Dockerfiles based on an
appropriate Rocker image. The Dockerfiles in the Rocker stack should
themselves provide a simple of example of this. A user begins by
selecting an appropriate base image for their needs: if the
RStudio\textsuperscript{\textregistered} interface is desired, a user
might start with \texttt{FROM\ rocker/rstudio}; an image for testing a
particular C code in an R package might use
\texttt{FROM\ rocker/r-devel-san}, an image for reproducing a data
analysis will probably select a stable version tag in addition to an
appropriate base library, \emph{e.g.},:
\texttt{FROM\ rocker/tidyverse:3.4.1} Users can easily add additional
software to any running Rocker image using the standard R and Debian
mechanisms. Details on how to extend Rocker images can be found at
\url{https://rocker-project.org}.

@@

The Rocker project consists of a suite of images built automatically by
and hosted on the Docker Hub, \url{https://hub.docker.com/r/rocker}.
Source Dockerfiles, supporting scripts and documentation are hosted on
GitHub under the organization \texttt{rocker-org},
\url{https://github.com/rocker-org}. The issue tracker and pull requests
are used for community input, discussions and contributions to these
images. The Rocker project wiki provides a place to synthesize
community-contributed documentation, use-cases, and other knowledge
about using the Rocker images.

@@

The \texttt{debian:testing} stack aims to make the most efficient use of
upstream builds: the pre-compiled \texttt{.deb} binaries provided by the
Debian repositories. It is both quicker and easier to install software
from binaries, since the package manager (\texttt{apt}) manages the
necessary (binary) dependencies and bypasses the time-consuming process
of compiling from source. Basing this stack on \texttt{debian:testing}
means that much more recent versions of commonly-used libraries and
compilers are available as binaries than would be found in a Debian
stable release. In order to provide access to the most recent available
binaries, this stack uses apt-pinning \citep{apt_pinning} to allow the
\texttt{apt} package manager to also install binaries from
\texttt{debian:unstable}, which represents the most recent, bleeding
edge of packages built for Debian when necessary. For instance, the
\texttt{r-base} image provided by Rocker installs the most recent
version of R as a binary from Debian \texttt{unstable}. Similarly,
recent versions of many popular R packages can also be installed through
the package manager, \emph{e.g.}, \texttt{apt-get\ install\ r-cran-xml}.
This can be particularly helpful for packages with external system
dependencies (such as \texttt{libxml2-dev} in this example) which cannot
be installed from the R console as they are system dependencies rather
than R packages installed from within R. We should note, however, that
only about 500 of the over 11,000 CRAN packages are available as Debian
packages.

As the names \texttt{testing} and \texttt{unstable} imply, this approach
is not without challenges. The particular version of any given package
can change as packages move from \texttt{unstable} into
\texttt{testing}. New versions are sent to \texttt{unstable} during the
normal course of Debian development. This can occasionally break an
previously-working installation command in a Dockerfile until the
maintainer redirects the package manager to install a package from the
\texttt{unstable} sources that could previously be installed from
\texttt{testing}, or vice versa (using the \texttt{-t} option in
\texttt{apt}). That said, packages only migrate from \texttt{unstable}
to \texttt{testing} after a period of several days---and if the
migration and installation of the particular version is free of
interactions with other packages in their dependency graph. That way,
\texttt{unstable} serves as validation lab which leaves \texttt{testing}
reasonably stable yet current.

Relative to \texttt{stable}, the \texttt{testing} stack thus offers the
following advantages:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  These Dockerfiles are easy to develop and extend because almost all
  software can be installed through the package manager.
\item
  These Dockerfiles always install the most recent available software.
\item
  These Dockerfiles can almost always build relatively quickly.
\end{enumerate}

and these down-sides:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  These Dockerfiles require occasional minor maintenance to ensure
  successful builds. (\emph{e.g.}, changing an installation directive
  from \texttt{unstable} to \texttt{testing} or vice versa).
\item
  The resulting images are inherently dynamic: rebuilding the same
  Dockerfile months or years apart will generate images with
  significantly different versions of software installed.
\item
  The use of apt pinning may be unfamiliar to some users, where the user
  must ultimately be responsible to ensure compatibility
  \citep{apt_pinning}.
\end{enumerate}

@@

The \texttt{debian:testing}-based stack currently includes seven images
actively maintained by the Rocker development team (Table 1).
\texttt{r-base} builds on \texttt{debian:testing}, and the other six in
the stack each build directly from \texttt{r-base}. The \texttt{r-base}
image is unique in that it is designated as the official image for the R
language by the Docker organization itself. This official image is
reviewed and then built by employees of Docker Inc based on a Dockerfile
maintained by the Rocker team. Consequently, users should refer to this
image in Docker commands without an organization namespace, \emph{e.g.},
\texttt{docker\ run\ -t\ r-base} to access the official image. All other
images in the Rocker project are not individually reviewed and built by
Docker Inc and must be referenced using the \texttt{rocker} namespace,
\emph{e.g.}, \texttt{docker\ run\ -ti\ rocker/r-devel}.

@@

This stack also includes the images \texttt{shiny} and
\texttt{rstudio:testing} that provide Shiny server and
RStudio\textsuperscript{\textregistered} server IDE from
RStudio\textsuperscript{\textregistered} Inc, built on the
\texttt{r-base} image. RStudio\textsuperscript{\textregistered} and
Shiny are registered trademarks of RStudio Inc, and their use and the
distribution of their software in binary form on Docker Hub has been
granted to the Rocker project by explicit permission from RStudio. Users
should review RStudio's trademark use policy
(\url{http://www.rstudio.com/about/trademark/}) and address inquiries
about further distribution or other questions to
\href{mailto:permissions@rstudio.com}{\nolinkurl{permissions@rstudio.com}}.
The Rocker project also provides images with
RStudio\textsuperscript{\textregistered} server and Shiny server in the
stable versioned stack.

@@

\textbf{MRAN archives}: To facilitate installation of only
contemporaneous versions of R packages on these images, the default CRAN
mirror from which to install R packages is fixed to a snapshot of CRAN
corresponding to the last date for which that version of R was the
latest release. These snapshots are provided by the MRAN archive created
by Revolution Analytics (now part of Microsoft). It archives daily
snapshots of all of CRAN from which a user can install packages with the
usual \texttt{install.packages()} function \citep{MRAN}. Users can
always override this default by passing any current CRAN repository
explicitly. Unlike CRAN, Bioconductor only updates its repositories
through bi-annual releases aligned to R's spring release schedule. Thus,
Bioconductor packages can be installed in the usual way using
\texttt{bioclite}, which automatically selects the Bioconductor release
corresponding to the version of R in use.

@@

The \texttt{tidyverse} image contains all required and suggested
dependencies of the commonly-used \texttt{tidyverse} and
\texttt{devtools} R packages, including external database libraries
(\emph{e.g.}, MariaDB and PostgreSQL). Users should consult the package
Dockerfiles or \texttt{installed.packages()} list directly for a
complete list of installed packages. The \texttt{verse} library adds
commonly-used dependencies, notably a large but not comprehensive LaTeX
environment and Java development libraries. Previously, the Rocker
project provided the image \texttt{hadleyverse} which was since divided
into \texttt{tidyverse} and \texttt{verse} through community input.

>>>>

Docker is a popular open-source tool to create, distribute, deploy, and
run software applications using \emph{containers}. Containers provide a
virtual environment (see \citet{Clark2014} for an overview of common
virtual environments) requiring all operating-system components an
application needs to run. Docker containers are lightweight as they
share the operating system kernel, start instantly using a layered
filesystem which minimizes disk footprint and download time, are built
on open standards that run on all major platforms (Linux, Mac, Windows),
and provide an added layer of security by running an application in an
isolated environment \citep{what-docker}. Familiarity with a few key
terms is helpful in understanding this paper. The term ``container''
refers to an isolated software environment on a computer. R users can
think of running a container as analogous to loading an R package; a
container is an active instance of a static Docker image. A Docker
``image'' is a binary archive of that software, analogous to an R binary
package: a given version is downloaded only once, and can then be
``run'' to create a container whenever it is needed. A ``Dockerfile'' is
a recipe, the source-code, to create a Docker image.\\
Pre-built Docker images are publicly available through Docker Hub, which
plays a role for central distribution similar to CRAN in our analogy.
Development and contributions to the Rocker project focus on the
construction, organization and maintenance of these Dockerfiles.

\section{Design principles and use
cases}\label{design-principles-and-use-cases}

Docker gives users very convenient access to pre-configured and
pre-built \emph{binary} images that ``just work''. This allows R users
to access a wider-variety of ready-to-use environments than provided by
either the R Project itself or, say, their distribution which will
generally focus on one (current) release. For example, R users on
Windows may run RStudio\textsuperscript{\textregistered} Server or
Shiny\textsuperscript{\textregistered} Server locally just by launching
a single command (once Docker itself is installed). Another common
use-case is access to R-devel without affecting the local system. Here,
we detail some of the principal use cases motivating these containerized
versions of R environments, and the design principles that help make
them work.

@@

One common use case for Rocker containers is to provide a fast and
reliable mechanism to deploy a custom R environment to a remote server,
such as Amazon Web Services Elastic Compute (AWS EC2), DigitalOcean,
NSF's Jetstream servers \citep{jetstream}, or private or institutional
server hardware. Rocker containers are also easy to run locally on most
modern laptops using Windows, MacOS, or Linux-based operating systems.
By sharing volumes with the local host, users can still manipulate files
with familiar, native tools while performing computation through a
reproducible, containerized environment \citep{Boettiger2015}. Being
able to test code in a predictable, pre-configured R environment on a
local machine and to then run the same code in an identical environment
on a remote server (\emph{e.g.}, for access to greater RAM, more
processors, or merely to free up the local machine from a long-running
computation) is essential for low-friction scaling of analysis. Without
such containerization, getting code to run appropriately in a remote
environment can be a major undertaking, requiring both time and
knowledge many would-be users may not have.

For instance, on any platform wiht Docker installed, the following
Docker command will launch a Rocker container providing the
rstudio-server environment over a web interface.

@@

The \texttt{docker\ run} option \texttt{-p} sets the port on which
RStudio\textsuperscript{\textregistered} will appear, for which 8787 is
the default (adding your user to the docker group to avoids the need for
a \texttt{sudo} command to call \texttt{docker}:
\texttt{sudo\ usermod\ -g\ docker\ \$USER}). Many academic and
commercial cloud providers make it possible to execute such code
snippets when a container is launched, without ever needing to
\texttt{ssh} into the machine. The user may log into the server merely
by pasting its IP address or DNS name (followed by the chosen port,
\emph{e.g.}, \texttt{:8787}) into a browser and entering the appropriate
password. This provides the user with a familiar, interactive
environment running on a remote machine while requiring a minimum of
expertise.

@@

\subsection{HPC application}\label{hpc-application}

The portability of Rocker images can be particularly valuable in High
Performance Computing contexts Setting up a specific R environment on
High Performance Computing platforms and other centrally administrated
multi-user machines or clusters has traditionally been challenging due
to restrictions on root access that may be needed to install certain
libraries. Versions of R and packages installed by the system
administrator may also lag behind the most recent releases. Deploying
Docker containers on HPC systems has previously been more very
problematic since most system adminstrators do not want to allow the
elevated user permissions the Docker runtime environment requires. To
work around this problem, Lawrence Berkeley National Labs (LBNL) has
made `Singularity' \citep{singularity}: a container runtime environment
that users can both install and use to run most Docker containers
without requiring root privileges. Singularity has seen rapid adoption
in the HPC community (\url{http://singularity.lbl.gov/install-request}).
Rocker containers can be run through Singularity with a single command
much like the native Docker commands, \emph{e.g.}

\begin{verbatim}
singularity exec docker://rocker/tidyverse:latest R
\end{verbatim}

More details can be found in the Singularity documentation.

@@

Users can access a \texttt{bash} shell running as root within a Rocker
container using

@@

As with any interactive Docker container, users should specify the
terminal (\texttt{-t}) and interactive (\texttt{-i}) flags, (here
combined with interactive as \texttt{-ti}), and specify the desired
executable environment (\emph{e.g.}, \texttt{R}, though other common
options may be \texttt{Rscript} or \texttt{bash}):

@@

Another feature of Rocker containers is the ability to provide a
sandboxed environment, isolated from software and potentially from other
data on the machine. Many users are reluctant to upgrade their suite of
installed packages, which may break their existing code or even their R
environment if the installation goes poorly. However, upgrading packages
and/or the R environment is often necessary to run analysis of a
colleague, or access more recent methods. Rocker offers an easy
solution. For instance, a user can run R code requiring the most recent
versions of R and related packages inside a Rocker container without
having to upgrade their local installations first. Conversely, one could
use Rocker to run code on an older R release with prior versions of R
packages, again without having to make any alteration to one's local R
install. Another common use case is to access a container with support
for particular options such as using gcc or clang compiler sanitizers
\citep{edd_sanitizers}. These require R itself be built with specialized
settings that may not be not available or familiar to many R users on
their native system, but can be easily deployed by pulling the Rocker
images \texttt{rocker/r-devel-san} or
\texttt{rocker/r-devel-ubsan-clang}.

This sandboxing feature is also valuable in the remote computing
context, allowing system administrators to grant users freedom to
install software which requires root privileges inside a container,
while not granting them root access on the host machine. Root access is
required to launch Docker containers, though not to access containers
already running and providing some service such as RStudio. Users
logging into a container through the
RStudio\textsuperscript{\textregistered} interface do not by default
have root privileges, though are able to install R packages. Granting
these users root privileges in the container still leaves them sandboxed
from the host container.\\
Sandboxing also serves an important function in reproducible research by
making it easier to test a specified environment in isolation from the
host machine. Unlike traditional virtual machines, these containers do
not impose a large footprint of reserved resources as a typical host can
easily support 100s of containers \citep{what-docker}.

@@

Users can easily determine the software stack installed on any Rocker
image by examining the associated Dockerfile recipe, which provides a
concise, human-readable record of the installation. All Rocker images
use automated builds through Docker Hub, which also acts as the central,
default repository distributing the images. Using automated builds
rather than uploading pre-built image binaries to Docker Hub avoids the
potential for the build not to match the recipe. The corresponding
Dockerfile is visible both on the Docker Hub and in the linked GitHub
repository, which provides a transparent versioned history of all
changes made to these recipes, as well as documentation, a community
wiki, and issue trackers for discussing proposed changes, bugs,
improvements to the Dockerfiles and troubleshoot any issues users may
encounter. Having these public source files built automatically by a
trusted provider (Docker Hub), rather than built locally and uploaded as
binaries, is also useful from a security perspective in avoiding
malware.

@@

Having a shared, transparent computational environment created by a
publicly hosted, reproducible recipe facilitates community input into
configuration details. R and many of its packages and related software
can be configured with a wide range of options, compilers, different
linear-algebra libraries and so forth. While this flexibility reflects
varying needs, many users rely on default settings which are most often
are optimized more for simplicity of installation rather than than
performance. The Rocker recipes reflect significant community input on
these choices. This helps create a more finely tuned, optimized
reference implementation of the R environment as well as a platform for
comparing and discussing these concerns which are often overlooked
elsewhere. Issues and Pull Requests on the Rocker repositories on GitHub
attest to some of these discussions and improvements. In particular,
input from the Docker Inc employees through the official approval
process for the \texttt{r-base} image, expertise from the Debian R
maintainer and other Debian developers, and both direct and indirect
feedback from the experience and user-generated documentation from many
early adopters in the R community has helped shape and strengthen the
project over the past few years. Widespread use of the Rocker image
helps promote both testing of these choices and contributions further
tweaking the configuration from many members of the R community.

@@

To make extensions transparent and persistent, Rocker images can be
extended by any user by writing their own Dockerfiles based on an
appropriate Rocker image. The Dockerfiles in the Rocker stack should
themselves provide a simple of example of this, (as described in the
following section).\\
A user begins by selecting an appropriate base image for their needs: if
the RStudio\textsuperscript{\textregistered} interface is desired, a
user might start with \texttt{FROM\ rocker/rstudio}; an image for
testing an R package with compiled code might use
\texttt{FROM\ rocker/r-devel-san}, an image for reproducing a data
analysis will probably select a stable version tag in addition to an
appropriate base library, \emph{e.g.},:
\texttt{FROM\ rocker/tidyverse:3.4.1} Users can easily add additional
software to any running Rocker image using the standard R and Debian
mechanisms. Details on how to extend Rocker images can be found at
\url{https://rocker-project.org}.

@@

The Rocker project consists of a suite of images built automatically by
and hosted on the Docker Hub, \url{https://hub.docker.com/r/rocker}.
Source Dockerfiles, supporting scripts and documentation are hosted on
GitHub under the organization \texttt{rocker-org},
\url{https://github.com/rocker-org}. The issue tracker and pull requests
are used for community input, discussions and contributions to these
images. The Rocker project wiki,
\url{https://github.com/rocker-org/rocker/wiki}, provides a place to
synthesize community-contributed documentation, use-cases, and other
knowledge about using the Rocker images.

@@

The \texttt{debian:testing} stack aims to make the most efficient use of
upstream builds: the pre-compiled \texttt{.deb} binaries provided by the
Debian repositories. It is both quicker and easier to install software
from binaries, since the package manager (\texttt{apt}) manages the
necessary (binary) dependencies and bypasses the time-consuming process
of compiling from source. Basing this stack on \texttt{debian:testing}
means that much more recent versions of commonly-used libraries and
compilers are available as binaries than would be found in a Debian
stable release. In order to provide \emph{optional} access to the most
recent available binaries, this stack uses apt-pinning
\citep{apt_pinning} to allow the \texttt{apt} package manager to
\emph{selectively} install binaries from \texttt{debian:unstable}, which
represents the most recent set of packages built for Debian. Similarly,
recent versions of many popular R packages can also be installed
pre-built through the package manager, \emph{e.g.},
\texttt{apt-get\ install\ r-cran-xml}. This can be particularly helpful
for packages with external system dependencies (such as
\texttt{libxml2-dev} in this example) which cannot be installed from the
R console as they are system dependencies rather than R packages
installed from within R. We should note, however, that only about 500 of
the over 11,000 CRAN packages are available as Debian packages.

As the names \texttt{testing} and \texttt{unstable} imply, particular
versions of package can change as packages move from \texttt{unstable}
into \texttt{testing}. New versions are sent to \texttt{unstable} during
the normal course of Debian development. This can occasionally break an
previously-working installation command in a Dockerfile until the
maintainer redirects the package manager to install a package from the
\texttt{unstable} sources that could previously be installed from
\texttt{testing}, or vice versa (using the \texttt{-t} option in
\texttt{apt}). That said, packages only migrate from \texttt{unstable}
to \texttt{testing} after a period of several days---and if the
migration and installation of the particular version is free of
interactions with other packages in their dependency graph. That way,
\texttt{unstable} serves as validation lab which leaves \texttt{testing}
reasonably stable yet current.

Relative to \texttt{stable}, the \texttt{testing} stack thus offers some
advantages as almost all software can be installed through the package
manager. Installation of binary packages from \texttt{testing} generally
provides the most recent available software, and installs it quickly as
a binary. On the other hand, these Dockerfiles may require occasional
maintenance when packages migrate and/or versions change. The resulting
images are also inherently dynamic: rebuilding the same Dockerfile
months or years apart will generate images with significantly different
versions of software installed as the pool of underlying packages
changes through time.

@@

The \texttt{debian:testing}-based stack currently includes seven images
actively maintained by the Rocker development team (Table 1).
\texttt{r-base} builds on \texttt{debian:testing}, and the other six in
the stack each build directly from \texttt{r-base}. The \texttt{r-base}
image is unique in that it is designated as the official image for the R
language by the Docker organization itself. This official image is
reviewed and then built by employees of Docker Inc based on a Dockerfile
maintained by the Rocker team. Consequently, users should refer to this
image in Docker commands without an organization namespace, \emph{e.g.},
\texttt{docker\ run\ -ti\ r-base} to access the official image. All
other images in the Rocker project are not individually reviewed and
built by Docker Inc and must be referenced using the \texttt{rocker}
namespace, \emph{e.g.}, \texttt{docker\ run\ -ti\ rocker/r-devel}.

@@

This stack also includes the images \texttt{shiny} and
\texttt{rstudio:testing} that provide Shiny server and
RStudio\textsuperscript{\textregistered} server IDE from
RStudio\textsuperscript{\textregistered} Inc, built on the
\texttt{r-base} image. RStudio\textsuperscript{\textregistered} and
Shiny are registered trademarks of RStudio Inc, and their use and the
distribution of their software in binary form on Docker Hub has been
granted to the Rocker project by explicit permission from RStudio. Users
should review RStudio\textsuperscript{\textregistered}'s trademark use
policy (\url{http://www.rstudio.com/about/trademark/}) and address
inquiries about further distribution or other questions to
\href{mailto:permissions@rstudio.com}{\nolinkurl{permissions@rstudio.com}}.
The Rocker project also provides images with
RStudio\textsuperscript{\textregistered} server and Shiny server in the
stable versioned stack.

@@

\textbf{MRAN archives}: To facilitate installation of only
contemporaneous versions of R packages on these images, the default CRAN
mirror from which to install R packages is fixed to a snapshot of CRAN
corresponding to the last date for which that version of R was current
(e.g. \texttt{3.4.2} was released on 2017-09-28, thus \texttt{3.4.1} is
pinned to the MRAN snapshot for that date). These snapshots are provided
by the MRAN archive created by Revolution Analytics (now part of
Microsoft). It archives daily snapshots of all of CRAN from which a user
can install packages with the usual \texttt{install.packages()} function
\citep{MRAN}. Users can always override this default by passing any
current CRAN repository explicitly. Unlike CRAN, Bioconductor only
updates its repositories through bi-annual releases aligned to R's
spring release schedule. Thus, Bioconductor packages can be installed in
the usual way using \texttt{bioclite}, which automatically selects the
Bioconductor release corresponding to the version of R in use.

@@

The \texttt{tidyverse} image contains all required and suggested
dependencies of the commonly-used \texttt{tidyverse} and
\texttt{devtools} R packages, including external database libraries
(\emph{e.g.}, MariaDB and PostgreSQL). Users should consult the package
Dockerfiles or \texttt{installed.packages()} list directly for a
complete list of installed packages. The \texttt{verse} library adds
commonly-used dependencies, notably a large but not comprehensive LaTeX
environment and Java development libraries. Previously, the Rocker
project provided the image \texttt{hadleyverse} which was since divided
into \texttt{tidyverse} and \texttt{verse} based on community input.
